\documentclass[usenames,dvipsnames]{article}
\usepackage{nips15submit_e,times}

\input{preamble/preamble.tex}
\input{preamble/preamble_acronyms.tex}
\input{preamble/preamble_math.tex}
\input{preamble/preamble_tikz.tex}

\setlength{\marginparwidth}{1in} % for margins for sidenotes
\usepackage{setspace}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{assumption}[theorem]{Assumption}

\title{Copula variational inference}

\author{
Dustin Tran\\
Harvard University\\
%\texttt{dtran@g.harvard.edu} \\
\And
David M. Blei \\
Columbia University \\
%\texttt{blei@cs.columbia.edu} \\
\And
Edoardo M. Airoldi  \\
Harvard University \\
%\texttt{airoldi@fas.harvard.edu} \\
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}
  We develop a general variational inference method that preserves
  dependency among the latent variables. Our method uses copulas to
  augment the families of distributions used in mean-field and
  structured approximations. Copulas model the dependency that is not
  captured by the original variational distribution, and thus the
  augmented variational family guarantees better approximations to the
  posterior. With stochastic optimization, inference on the augmented
  distribution is scalable. Furthermore, our strategy is generic: it
  can be applied to any inference procedure that currently uses the
  mean-field or structured approach.
  Copula variational inference has many advantages: it reduces bias;
  it is less sensitive to local optima; it is less sensitive to
  hyperparameters; and it helps characterize and interpret the
  dependency among the latent variables.
\end{abstract}

\input{sec_introduction.tex}
\input{sec_background.tex}
\input{sec_cvi.tex}
\input{sec_experiments.tex}
\input{sec_conclusion.tex}

\subsubsection*{Acknowledgments}
We thank Luke Bornn, Robin Gong, and Alp Kucukelbir for their
insightful comments.
This work is supported by NSF IIS-0745520, IIS-1247664, IIS-1009542,
ONR N00014-11-1-0651, DARPA FA8750-14-2-0009, N66001-15-C-4032,
Facebook, Adobe, Amazon, and the John Templeton Foundation.

\bibliographystyle{apalike}
\small
\bibliography{nips2015}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
